Microxanox - an R package to simulate ...

### 2022-03-21 inst/manuscript/manuscript_microxanox.Rmd
Next edits

---
title: "m0.0.1: Microxanox - ..."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: | 
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.
journal: 'Ecology and Evolution - Applications'
bibliography: manuscript_bibliography.bib
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{setspace}
- \doublespacing
- \usepackage{fancyhdr}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[CO,CE]{Metadata made easy - the dmdScheme approach}
- \pagestyle{fancy}

nocite: '@Bush2017'
linkcolor: blue
link-citations: yes
csl: methods-in-ecology-and-evolution.csl
---

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. {>>If we wnat that, we havt o include some, even rudimentary, GITHUB tests.<<}

@@

# Using the package

@@

## The simulation framework

Here make an overview of how conditions and parameters are specified, stored, and results stored. I think there is text in the user guide vignette that can be used (duplicated) here.

We will discuss the general structure and functiuonalit of the package without going into to mu=ch detail.

@@

In the following sections we describe the general goals of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.

## Running one simulation

@@

## Find a Steady State of the model

@@

### Replication Method
The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.

### Temporal Method
The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step. We hereafter term this "the temporal method".
This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .

@@

## Visualisation

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in Figure 1.

**Please include figures so they can have captions and be cross referenced, as in the Supplement.**

![](./figures/fig1_plot_dynamics.png)

@@

The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.

@@

Take from the user guide.

@@

Take from the partial reproduction.

@@

This package has been used in the production of the paper [@REF_NEEDED]. It can be used for similar studies in this study systems. It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs. Consequently, any more substantial, are likely to need a change in the source code.

Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g. the event definition can easily be changed. or other aspects can be adjusted. All values in the parameter object can be changed.

@@

>>>>

---
title: "Microxanox - an R package to simulate ..."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: | 
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.
bibliography: manuscript_bibliography.bib
date: "`r Sys.time()`"

# RMarkdown output settings
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
  bookdown::pdf_document2:
    keep_tex: yes

# qarto output settings
format:
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    link-external-icon: true
    title-block-style: default

# filters:
#   - type: json
#     path: criticpandoc.py

# Citation settings    
nocite: '@Bush2017'
linkcolor: blue
link-citations: yes

csl: methods-in-ecology-and-evolution.csl
---

@@

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
{>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}

@@

# The R Package

@@

## The modelling framework

The framework used when writing this package aims at reproducibility of the results. It builds on the folowing main considerations:

1. all parameter needed to run a simulation or find a stable state are contained in a single parameter object. This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`. Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim)  and in the [User Guide](LINK_NEEDED).
2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additionl slot named `results` which contains the results of the run
4. As this return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.

The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.

A typical simulation would look as followed:

```{r, eval = FALSE}
## Create the parameter
parameter <- new_runsim_parameter()
# manually setting certain parameter

## Run the simulation and save the result
result <- run_simulatrion(parameter)
saveRDS(result, "sim1.rds")

## Do other stuff, e.g. plotting

## Load results, change some parameter, and rerun the simulation and save the result
parameter <- loadRDS("sim1.rds")
# change some parameter
result <- run_simulatrion(parameter)
saveRDS(result, "sim2,rds")
```

## Using the package {#runsim}

We will now discuss the general structure and functiuonality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).

@@

In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.

### Running one simulation

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).

```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function `plot_dynamics()`$`. Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
knitr::include_graphics("figures/fig1_plot_dynamics.png")
```

### Find a Steady State of the model

@@

#### Replication Method
The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.

#### Temporal Method
The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.

This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .

@@

@@

The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package. All of these use cases can be expanded to larger numbers of strains per functional group.

@@

In the [User Guide](LINK_NEEDED) we ued a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to analyse as an example the determination of the stable states during temporal environmental chan ge (the oxygen diffusivity). From these simulatiuons, we extracted measures of nonlinearity and hysteresis.

@@

One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED). All aspects in the paper could be reproduced.

@@

This model (as part of the package) has been used in the production of the paper [@REF_NEEDED]. It can be used for similar studies in this study systems.

{>>More details from the paper.<<}

## General usability and flexibility of the model
The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED]. Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.

Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g. the event definition can easily be changed. or other aspects can be adjusted. All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.

@@

{>>Dependant on journal<<}
